---
title: "CNV caller report"
author: "Stef Piatek"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
params:
  capture: "multi"
  
---

Script to do a quick analysis of specific capture

Example usage from command line:

    Rscript -e "rmarkdown::render('cnv-caller-report.Rmd', params = list(capture = 'ICR'), 'html_document', 'output/ICR_report.html')"


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(here)
library(broom)
library(purrr)
library(glue)
library(ggsci)

capture_is_icr <- params$capture == "ICR"
```

## Parse database to get summary of all output

Connecting to database and grabbing out known positives, known negatives, true positives and false negatives


```{r database}
source(here::here("helper_functions", "database.R"))
# note filtering out cnv 37 from minipanel for BRCA1 and CNV 115 for BRCA2. GATK called in all samples


if (params$capture != "multi"){
  database_path <- here("../../output", glue("{params$capture}.sqlite"))
  
  true_positives <- get_true_positives(database_path )
  positives <- get_positives(database_path)
  false_negatives <- get_false_negatives(database_path) 
  negatives <- get_negatives(database_path)
  callers <- get_callers_list(database_path) %>%
    pull(name)
} else {
  captures <- here("../../output", paste0(c("ICR_n2", "MiniPanel", "Cancer_v1"), ".sqlite"))
  # purr
  
  true_positives <- map_dfr(.x = captures, .f = get_true_positives) 
  positives <- map_dfr(.x = captures, .f = get_positives)
  false_negatives <- map_dfr(.x = captures, .f = get_false_negatives) 
  negatives <- map_dfr(.x = captures, .f = get_negatives)
  callers <- map_df(.x = captures, .f = get_callers_list) %>%
    pull(name) %>%
    as.factor() %>%
    levels()

}




```

## Caller analysis

Note: Hidden likeihood ratio calculation 

```{r def_likelihood, echo=FALSE}
source(here::here("helper_functions", "likelihood_ratio.R"))

```

Analyse calls and give upper and lower 95% confidence intervals


```{r analyse_calls, results='asis'}

get_binom_ci <- function(called_num, total_num, type_name, caller_name){
  prop.test(called_num, total_num) %>%
    tidy() %>%
    mutate(caller = caller_name,
           total = total_num,
           called =  called_num,
           type = type_name) %>%
    select(caller, type, estimate, conf.low, conf.high, called, total, everything()) %>%
    return()
}


analyse_caller <- function(caller_name){
  caller_false_negatives <- false_negatives %>%
    filter(name_caller == caller_name) 
  
  caller_true_positives <- true_positives %>%
    filter(name_caller == caller_name) 
  
  # statuses
  true_pos <- nrow(caller_true_positives)
  true_neg <- nrow(negatives) - nrow(caller_false_negatives)
  false_pos <- nrow(negatives) - true_neg
  false_neg <- nrow(positives) - true_pos

  total_stats <- rbind(
    get_binom_ci(true_pos, true_pos + false_neg, "sensitivity", caller_name),
    get_binom_ci(true_neg, false_pos + true_neg, "specificity", caller_name),
    get_binom_ci(true_pos + true_neg, true_pos + false_neg + false_pos + true_neg, "accuracy", caller_name),
    ## Must fit with the actual diseas prevalence so not used
    # get_binom_ci(true_pos, true_pos + false_pos, "ppv", caller_name),
    # get_binom_ci(true_neg, false_neg + true_neg, "npv", caller_name),
    likelihood_ratio(true_pos, false_pos, false_neg,  true_neg, "plr", caller_name),
    likelihood_ratio(true_pos, false_pos, false_neg,  true_neg, "nlr", caller_name)
    )
  return(total_stats)
}


caller_stats <- map_dfr(callers, analyse_caller) %>%
  mutate(type = str_replace(type, "nlr", "Negative Likelihood Ratio")) %>%
  mutate(type = str_replace(type, "plr", "Positive  Likelihood Ratio")) %>%
  mutate(type = str_to_title(type))

caller_stats %>%
  filter(type %in% c("Accuracy", "Sensitivity", "Specificity")) %>%
  ggplot(aes(x=caller, colour=caller)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high)) +
  geom_point(aes(y=estimate)) +
  facet_wrap(~type) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  ggsci::scale_colour_jco() +
  labs(x = "CNV Caller",
       y = "Value") +
  guides(colour=FALSE) + 
  ggsave(here::here("../../output/plots_tables/", glue("{params$capture}_all_caller_metrics.pdf")), height = 5) +
  ggsave(here::here("../../output/plots_tables/", glue("{params$capture}_all_caller_metrics.png")), height = 5)

caller_stats %>%
  filter(!type %in% c("Accuracy", "Sensitivity", "Specificity")) %>%
  # filter(!caller %in% c("copywriter", "xhmm")) %>%
  ggplot(aes(x=caller, colour=caller)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high)) +
  geom_point(aes(y=estimate)) +
  facet_wrap(~type, scales = "free_y") +
  labs(x = "CNV Caller",
       y = "Value") +
  guides(colour=FALSE) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  ggsci::scale_colour_jco()  +
  ggsave(here::here("../../output/plots_tables/", glue("{params$capture}_caller_likelihood_ratios.pdf")), height = 4, width = 4) +
  ggsave(here::here("../../output/plots_tables/", glue("{params$capture}_caller_likelihood_ratios.png")), height = 4, width = 4)


caller_stats %>%
  select(caller:total) %>%
  filter(type %in% c("Sensitivity", "Specificity", "Accuracy")) %>%
  knitr::kable() %>%
  print()
  
caller_stats %>%
  select(caller:conf.high) %>%
  filter(!type %in% c("Sensitivity", "Specificity", "Accuracy")) %>%
  knitr::kable() %>%
  print()


# decon vs exome depth
false_negatives %>%
  rbind(true_positives) %>%
  filter(name_caller %in% c("decon", "exome-depth")) %>%
  full_join(positives, by = c("sample_id" = "id", "result_type"), suffix = c("_unknown", "")) %>%
  group_by(sample_id) %>%
  filter(n() < 2,
         result_type == "positive") %>%
  mutate(uncalled = is.na(caller_id)) %>%
  arrange(uncalled, sample_id) %>%
  ungroup() %>% 
  rename(sample_name = name,
         caller = name_caller,
         gene = name_gene) %>%
  arrange(capture, uncalled, gene, sample_name) %>%
  select(gene, sample_name, caller, capture) %>% 
  write_csv(here::here(glue("../../output/plots_tables/{params$capture}_exon_exome.csv")), na = "Not Detected")

# top4 missed
false_negatives %>%
  rbind(true_positives) %>%
  filter(name_caller %in% c("decon", "exome-depth", "gatk", "savvycnv")) %>%
  full_join(positives, by = c("sample_id" = "id", "name_gene", "capture", "name", "result_type"), 
            suffix = c("_unknown", "")) %>%
  mutate(uncalled = is.na(caller_id)) %>%
  group_by(sample_id, capture, name_gene) %>%
  filter(result_type == "positive") %>%
  filter(n() < 4) %>%
  rename(sample_name = name,
         caller = name_caller,
         gene = name_gene) %>%
  summarize(caller = toString(caller),
            uncalled = min(uncalled),
            sample_name = min(sample_name)) %>%
  arrange(capture, gene, uncalled, sample_name) %>%
  select(gene, sample_name, caller, capture) %>% 
  write_csv(here::here(glue("../../output/plots_tables/{params$capture}_total_missing.csv")), na = "Not Detected")

## venn diagram
library(VennDiagram)

top4 <- true_positives %>%
  filter(name_caller %in% c("decon", "exome-depth", "gatk", "savvycnv")) %>%
  mutate(name_gene = paste0(name, name_gene)) # honestly didn't expect two CNVs in different genes but there we go

venn.diagram(x = list(
    "DECoN" = top4 %>% filter(name_caller == "decon") %>% pull(name_gene),
    "GATK" = top4 %>% filter(name_caller == "gatk") %>% pull(name_gene),
    "SavvyCNV" = top4 %>% filter(name_caller == "savvycnv") %>% pull(name_gene),
    "Exome Depth" = top4 %>% filter(name_caller == "exome-depth") %>% pull(name_gene)
    ),
    fill = ggsci::pal_jco()(4), 
    filename = here::here("../../output/plots_tables/", "top4_callers.png"),
    resolution = 600,
    imagetype = "png",
    height = 4,
    width = 4,
    units = "in"
)

```

Determine threshold for copywriter if ICR data is being used:

0.5 threshold for calling CNVs


```{r copywriter, eval=capture_is_icr}

library(plotROC)
library(jsonlite)


get_seg_mean <- function(json_data){
  # returns absolute value of seg.mean for copywriter
  json <- parse_json(json_data)
  seg_mean <- abs(json$seg.mean)
  return(seg_mean)
}

copywriter_data <- false_negatives %>%
  rbind(true_positives) %>%
  filter(caller_id == 2) %>%
  rowwise() %>%
  mutate(seg_mean = get_seg_mean(json_data)) %>%
  full_join(positives, by = c("sample_id" = "id", "result_type")) %>%
  full_join(negatives, by = c("sample_id" = "id", "result_type")) %>%
  mutate(roc_outcome = if_else(result_type == "normal", 0, 1),
         seg_mean = if_else(is.na(seg_mean), 0, seg_mean)) 

ggplot(copywriter_data, aes(d = roc_outcome, m = seg_mean)) +
  geom_abline(slope=1, intercept=0, colour="grey") +
geom_roc(linealpha = 0.3, colour = "#005aac", cutoffs.at = c(0:10 * 0.1, 2)) +
  theme_bw() +
  ylab("Sensitivity") +
  xlab("1 - Specificity")
save.image(here("../../output", params$capture, "copywriter_threshold.RData"))

```

```{r save_data}

save.image(here("../../output", glue("{params$capture}caller_analysis.RData")))
```

