---
title: "CNV caller report"
author: "Stef Piatek"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
     keep_md: TRUE
params:
  capture: "ICR"
  
---

Script to do a quick analysis of specific capture

Example usage from command line:

    Rscript -e "rmarkdown::render('cnv-caller-report.Rmd', params = list(capture = 'ICR'), 'html_document', 'output/ICR_report.html')"


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DBI)
library(RSQLite)
library(tidyverse)
library(knitr)
library(here)
library(broom)
library(purrr)
library(glue)
```

## Parse database to get summary of all output

Connecting to database and grabbing out known positives, known negatives, true positives and false negatives


```{r database}
con <- dbConnect(SQLite(), dbname= here("output", "results.sqlite"))

genes <- con %>%
  tbl("genes") %>%
  filter(capture == params$capture) 

positives <- genes %>%
  left_join(tbl(con, "samples"), by = c("id" = "gene_id"), suffix = c("_gene_join", "")) %>%
  select(-ends_with("_gene_join")) %>%
  filter(result_type == "positive")

true_positives <- positives %>%
  left_join(tbl(con, "called_cnvs"), by=c("id" = "sample_id"),  suffix = c("", "_called_cnv")) %>%
  collect() %>%
  mutate(sample_id = id) %>%
  select(-id)

positives <- positives %>%
  collect()

# set this up to instead be a list from the run. do this later though
negatives <-  genes %>%
  left_join(tbl(con, "samples"), by = c("id" = "gene_id"), suffix = c("_gene_join", "")) %>%
  select(-ends_with("_gene_join")) %>%
  filter(result_type == "normal") 

false_negatives <- negatives %>%
  left_join(tbl(con, "called_cnvs"), by=c("id" = "sample_id"),  suffix = c("", "_called_cnv")) %>%
  collect() %>%
  mutate(sample_id = id) %>%
  select(-id)

negatives <- negatives %>%
  collect() 

callers <- con %>%
  tbl("callers") %>%
  collect()

dbDisconnect(con)
```

## Caller analysis

Note: Hidden likeihood ratio calculation 

```{r def_likelihood, echo=FALSE}
likelihood_ratio <- function(a, b, c, d, input_type="plr", input_caller) {
  # taken from https://stats.stackexchange.com/questions/61349/how-to-calculate-the-confidence-intervals-for-likelihood-ratios-from-a-2x2-table
  # 
  alpha <- 0.05

  spec <- d/(b+d)
  sens <- a/(a+c)

  lr.pos <- sens/(1 - spec)  

  if ( a != 0 & b != 0 ) {

    sigma2 <- (1/a) - (1/(a+c)) + (1/b) - (1/(b+d))
    lower.pos <- lr.pos * exp(-qnorm(1-(alpha/2))*sqrt(sigma2))
    upper.pos <- lr.pos * exp(qnorm(1-(alpha/2))*sqrt(sigma2)) 


  } else if ( a == 0 & b == 0 ) {

    lower.pos <- 0
    upper.pos <- Inf

  } else if ( a == 0 & b != 0 ) {

    a.temp <- (1/2)

    spec.temp <- d/(b+d)
    sens.temp <- a.temp/(a+c)
    lr.pos.temp <- sens.temp/(1 - spec.temp)  
    lower.pos <- 0
    sigma2 <- (1/a.temp) - (1/(a.temp+c)) + (1/b) - (1/(b+d))
    upper.pos <- lr.pos.temp * exp(qnorm(1-(alpha/2))*sqrt(sigma2))

  } else if ( a != 0 & b == 0 ) {

    b.temp <- (1/2)
    spec.temp <- d/(b.temp+d)
    sens.temp <- a/(a+c)
    lr.pos.temp <- sens.temp/(1 - spec.temp) 
    sigma2 <- (1/a) - (1/(a+c)) + (1/b.temp) - (1/(b.temp+d))
    lower.pos <- lr.pos.temp * exp(-qnorm(1-(alpha/2))*sqrt(sigma2))
    upper.pos <- Inf  

  } else if ( (a == (a+c)) & (b == (b+d)) ) {

    a.temp <- a - (1/2)
    b.temp <- b - (1/2)
    spec.temp <- d/(b.temp+d)
    sens.temp <- a.temp/(a+c)
    lr.pos.temp <- sens.temp/(1 - spec.temp) 
    sigma2 <- (1/a.temp) - (1/(a.temp+c)) + (1/b.temp) - (1/(b.temp+d))
    lower.pos <- lr.pos.temp * exp(-qnorm(1-(alpha/2))*sqrt(sigma2))
    upper.pos <- lr.pos.temp * exp(qnorm(1-(alpha/2))*sqrt(sigma2)) 

  }

  lr.neg <- (1 - sens)/spec

  if ( c != 0 & d != 0 ) {

    sigma2 <- (1/c) - (1/(a+c)) + (1/d) - (1/(b+d))
    lower.neg <- lr.neg * exp(-qnorm(1-(alpha/2))*sqrt(sigma2))
    upper.neg <- lr.neg * exp(qnorm(1-(alpha/2))*sqrt(sigma2)) 

  } else if ( c == 0 & d == 0 ) {

    lower.neg<- 0
    upper.neg <- Inf

  } else if ( c == 0 & d != 0 ) {

    c.temp <- (1/2)

    spec.temp <- d/(b+d)
    sens.temp <- a/(a+c.temp)
    lr.neg.temp <- (1 - sens.temp)/spec.temp    
    lower.neg <- 0
    sigma2 <- (1/c.temp) - (1/(a+c)) + (1/d) - (1/(b+d))
    upper.neg <- lr.neg.temp * exp(qnorm(1-(alpha/2))*sqrt(sigma2))

  } else if ( c != 0 & d == 0 ) {

    d.temp <- (1/2)
    spec.temp <- d.temp/(b+d)
    sens.temp <- a/(a+c)
    lr.neg.temp <- (1 - sens.temp)/spec.temp  
    sigma2 <- (1/c) - (1/(a+c)) + (1/d.temp) - (1/(b+d))
    lower.neg <- lr.neg.temp * exp(-qnorm(1-(alpha/2))*sqrt(sigma2))
    upper.neg <- Inf  

  } else if ( (c == (a+c)) & (d == (b+d)) ) {

    c.temp <- c - (1/2)
    d.temp <- d - (1/2)
    spec.temp <- d.temp/(b+d)
    sens.temp <- a/(a+c.temp)
    lr.neg.temp <- (1 - sens.temp)/spec.temp   
    sigma2 <- (1/c.temp) - (1/(a+c)) + (1/d.temp) - (1/(b+d))
    lower.neg <- lr.neg.temp * exp(-qnorm(1-(alpha/2))*sqrt(sigma2))
    upper.neg <- lr.neg.temp * exp(qnorm(1-(alpha/2))*sqrt(sigma2)) 

  }
  # Edited code from here
  if (input_type == "plr"){
    ratio <- lr.pos
    lower <- lower.pos
    upper <- upper.pos
  } else if (input_type == "nlr"){
      ratio <- lr.neg
    lower <- lower.neg
    upper <- upper.neg
  }
  
  output <- tibble(
      caller = input_caller,
      type = input_type,
      estimate = ratio,
      conf.low = lower,
      conf.high = upper,
      called = NA,
      total = NA,
      statistic = NA,
      p.value = NA,
      parameter = NA,
      method = NA,
      alternative = NA
      )
    return(output)
}

```

Analyse calls and give upper and lower 95% confidence intervals


```{r analyse_calls, results='asis'}

get_binom_ci <- function(called_num, total_num, type_name, caller_name){
  prop.test(called_num, total_num) %>%
    tidy() %>%
    mutate(caller = caller_name,
           total = total_num,
           called =  called_num,
           type = type_name) %>%
    select(caller, type, estimate, conf.low, conf.high, called, total, everything()) %>%
    return()
}

analyse_caller <- function(input_caller_id){
  caller_name <- callers %>%
    filter(id == input_caller_id) %>%
    pull(name)

  caller_false_negatives <- false_negatives %>%
    filter(caller_id == input_caller_id)
  
  caller_true_positives <- true_positives %>%
    filter(caller_id == input_caller_id)
  
  # statuses
  true_pos <- nrow(caller_true_positives)
  true_neg <- nrow(negatives) - nrow(caller_false_negatives)
  false_pos <- nrow(negatives) - true_neg
  false_neg <- nrow(positives) - true_pos

  total_stats <- rbind(
    get_binom_ci(true_pos, true_pos + false_neg, "sensitivity", caller_name),
    get_binom_ci(true_neg, false_pos + true_neg, "specificity", caller_name),
    get_binom_ci(true_pos + true_neg, true_pos + false_neg + false_pos + true_neg, "accuracy", caller_name),
    ## Must fit with the actual diseas prevalence so not used
    # get_binom_ci(true_pos, true_pos + false_pos, "ppv", caller_name),
    # get_binom_ci(true_neg, false_neg + true_neg, "npv", caller_name),
    likelihood_ratio(true_pos, false_pos, false_neg,  true_neg, "plr", caller_name),
    likelihood_ratio(true_pos, false_pos, false_neg,  true_neg, "nlr", caller_name)
    )
  return(total_stats)
}

caller_stats <- map_dfr(callers$id, analyse_caller)



caller_stats %>%
  select(caller:total) %>%
  filter(type %in% c("sensitivity", "specificity", "accuracy")) %>%
  knitr::kable() %>%
  print()
  
caller_stats %>%
  select(caller:conf.high) %>%
  filter(type %in% c("plr", "nlr")) %>%
  knitr::kable() %>%
  print()

```

Determine threshold for copywriter using ICR data:

0.5 threshold for calling CNVs


```{r copywriter}

library(plotROC)
library(jsonlite)


get_seg_mean <- function(json_data){
  # returns absolute value of seg.mean for copywriter
  json <- parse_json(json_data)
  seg_mean <- abs(json$seg.mean)
  return(seg_mean)
}

copywriter_data <- false_negatives %>%
  rbind(true_positives) %>%
  filter(caller_id == 2) %>%
  rowwise() %>%
  mutate(seg_mean = get_seg_mean(json_data)) %>%
  full_join(positives, by = c("sample_id" = "id", "result_type")) %>%
  full_join(negatives, by = c("sample_id" = "id", "result_type")) %>%
  mutate(roc_outcome = if_else(result_type == "normal", 0, 1),
         seg_mean = if_else(is.na(seg_mean), 0, seg_mean))

ggplot(copywriter_data, aes(d = roc_outcome, m = seg_mean)) +
  geom_abline(slope=1, intercept=0, colour="grey") +
  geom_roc(linealpha = 0.3, colour = "#005aac", cutoffs.at = c(0:10 * 0.1, 2)) +
  theme_bw() +
  ylab("Sensitivity") +
  xlab("1 - Specificity") 
```

